Going from coarse landings data to fine scale species distribution

### 2022-04-26 Version_M2_plusieurszones/draft/paper_3/paper_3.Rmd
remove generated files from git

output            :
  papaja::apa6_pdf
    # reference_docx: mystyles_number_test.docx
---

@@

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)

@@

# Material and methods

We base our approach on the model developped by @alglave_combining_2022. In this framework, all observations were supposed to be realised at a fine scale. Here, we extend the framework to potentially provide fine scale predictions from data that have a coarse resolution (landings data) and combine these data with other fine scale datasets such as scientific data.

## Defining the problem from a modelling point of view

**Definition of the variables**

Let's define the latent field $S$ and the punctual observations $Y$. $S$ is a spatial log-Gaussian Field (GF) defined as $\log(S(x))=\mu+\beta . \Gamma(x) + \delta(x)$ (Figure \@ref(fig:SchemeRealloc)). We consider a discrete study domain where $x \in [\![1,n]\!]$ stands for the discrete locations. $\beta . \Gamma(x)$ is the covariate term with $\Gamma$ the matrix of design of the covariates and $\beta$ the effect of the covariates. $\delta(x)$ is a GF capturing spatial correlation ($\delta \sim \mathcal{N}(0,\Sigma)$).

>>>>

output            :
  papaja::apa6_pdf:
    includes:
            in_header: header.sty
    # reference_docx: mystyles_number_test.docx
---

@@

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed, cache = TRUE)

@@

#  Introduction 
* Commencer par la motivation peche
--> knowledge of essential habitat --> Species distribution model --> link between biomass and habitat to identify essential habitat
-->  Data from scientific survey, precisely geolocalized standard sampling protocol, expensive and time consuming
--> On the other hand declaration data, spatialized at coarser level  but cheap and reglementaire
Combining the two sources of data.

This problem is not unique, other exampl in terms of apllication

Change of support form fine to coarse sptial level

A solution the reallocation process

** Pour la discu, simulation conditionnelle ? **

We proposed an alternative approach to account for the change of support and we compare it to the approach proposed in alglave but also to a golden standard to assess the consequences of these methods.

Paper organized as follow

Section 2. presentation of the reference model and detail the reallocation process

Section 3. Presnetation of an alternaion way to deal with change of support

Section 4. simulation study to asses the bias and/or the loss of precision of the two methodes compared with the golden standard

Section 5. Appllication on sole of the Bay of Biscay

Section 6. Discu/conclu

## An explicitly spatialized catch model

In  @alglave_combining_2022, the authors propose a hierarchical spatial model to combine scientific survey data obtained through a standardized sampling protocol and catch data harvested by fishermen. This section provides the key ingredient of this model and raises some concern on the spatialization of the commercial catch data.

Let  $D\subset \R^2$ be a spatial domain and $\bm{S}=(S(x), x\in D)$ the spatial repartition of the biomass for a species of interest.  $\bm{S}$ is assumed  to be a spatial log-Gaussian Random Field (GRF) defined as $\log(S(x))=\mu + \beta . \Gamma(x) + \delta(x)$ (Figure \@ref(fig:SchemeRealloc)) where $\bm{\delta}=(\delta(x), x\in D)$ is a zero mean GRF with covariance matrix $\Sigma$.  As the catch are   zero-inflated positive continuous data,  following  @thorson_three_2018,  the authors model a catch at a given site $X_i$ by a mixture model between a Dirac mass at $0$ and Log Normal distribution

\begin{equation}
Y_i | S(x_i) \overset{ind}{\sim} p_i \delta_0+ (1-p_i) B_i, \quad B_i \sim \mathcal{L}\left ( \frac{S(x_i)}{1-p_i} ,\sigma^2\right),
\end{equation}

<!-- 
Je pense qu'il ne faut pas parler de la grille discrete tout de suite car S est défini sur un espace continu et moralement l'intensité du processus de poisson aussi. le passage à la grille est la pour simplifier l'expression de la vraisemblance, on peut dire qu'on fait une approximation pour des raisons numériques

We consider a discrete study domain where $x \in [\![1,n]\!]$ stands for the discrete locations. $\beta . \Gamma(x)$ is the covariate term with $\Gamma$ the matrix of design of the covariates and $\beta$ the effect of the covariates. $\delta(x)$ is a GF capturing spatial correlation ($\delta \sim \mathcal{N}(0,\Sigma)$). -->
